A Formalization of Typed Lua

### 2015-08-21 doc/papers/dls15/paper.tex
camera-ready version for dls 2015

@@

\conferenceinfo{DLS '15}{October 25-30, 2015, Pittsburgh, PA, USA}
\copyrightyear{2015}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\doi{nnnnnnn.nnnnnnn}

@@

\authorinfo{André Murbach Maidl}
           {PUC-Rio}
           {amaidl@inf.puc-rio.br}
\authorinfo{Fabio Mascarenhas}
           {UFRJ}
           {fabiom@dcc.ufrj.br}
\authorinfo{Roberto Ierusalimschy}
           {PUC-Rio}
           {roberto@inf.puc-rio.br}

@@

\begin{figure}[t]
{\scriptsize
$$
\begin{array}{rlr}
F ::= & & \textsc{first-level types:}\\
& \;\; L & \textit{literal types}\\
& | \; B & \textit{base types}\\
& | \; \Nil & \textit{nil type}\\
& | \; \Value & \textit{top type}\\
& | \; \Any & \textit{dynamic type}\\
& | \; \Self & \textit{self type}\\
& | \; F_{1} \cup F_{2} & \textit{union types}\\
& | \; S_{1} \rightarrow S_{2} & \textit{function types}\\
& | \; \{\overline{F:V}\}_{unique|open|fixed|closed} & \textit{table types}\\
& |\; x & \textit{type variables}\\
& | \; \mu x.F & \textit{recursive types}\\
%\multicolumn{3}{c}{}\\
L ::= & & \textsc{{\tiny literal types:}}\\
&  \False \; | \; \True \; | \; {\it int} \; | \; {\it float} \; | \; {\it string} &\\
%\multicolumn{3}{c}{}\\
B ::= & & \textsc{{\tiny base types:}}\\
& \Boolean \ |\  \Integer \ |\ \Number\ |\ \String &\\\\
%\multicolumn{3}{c}{}\\
%\multicolumn{3}{c}{}\\
V ::= & & \textsc{{\tiny value types:}}\\
& \;\; F \; | \; \Const \; F &\\
S ::= & & \textsc{second-level types:}\\
& \;\; P & \textit{tuple types}\\
& | \; S_{1} \sqcup S_{2} & \textit{unions of tuple types}\\
%\multicolumn{3}{c}{}\\
P ::= & & \textsc{{\tiny tuple types:}}\\
%& \;\; \Void & \textit{void type}\\
& | \; F{*} & \textit{variadic types}\\
& | \; F \times P & \textit{pair types}\\\\
\end{array}
$$
}
\caption{Type Language}
\label{fig:typelang}
\end{figure}

@@

In our example, $x$ would be bound to $\Integer \times \Integer \times \Nil{*}$ in the {\bf then} branch, and would be bound
to $\Nil \times \String \times \Nil{*}$ inside the {\bf else}
branch. Now $q$ and $r$ project to $\Integer$ and $\Integer$
in the {\bf then} branch, and to $\Nil$ and $\String$ in the
{\bf else} branch.

>>>>

\toappear{}

@@

\conferenceinfo{DLS '15}{October 27th, 2015, Pittsburgh, PA, USA}
\copyrightyear{2015}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\doi{nnnnnnn.nnnnnnn}

@@

\authorinfo{André Murbach Maidl}
           {PUC-Rio}
           {amaidl@inf.puc-rio.br}
\authorinfo{Fabio Mascarenhas}
           {UFRJ}
           {mascarenhas@acm.org}
\authorinfo{Roberto Ierusalimschy}
           {PUC-Rio}
           {roberto@inf.puc-rio.br}

@@

\begin{figure}[t]
{\scriptsize
$$
\begin{array}{rlr}
F ::= & & \textsc{first-level types:}\\
& \;\; L & \textit{literal types}\\
& | \; B & \textit{base types}\\
& | \; \Nil & \textit{nil type}\\
& | \; \Value & \textit{top type}\\
& | \; \Any & \textit{dynamic type}\\
& | \; \Self & \textit{self type}\\
& | \; F_{1} \cup F_{2} & \textit{union types}\\
& | \; S_{1} \rightarrow S_{2} & \textit{function types}\\
& | \; \{\overline{F:V}\}_{unique|open|fixed|closed} & \textit{table types}\\
& |\; x & \textit{type variables}\\
& | \; \mu x.F & \textit{recursive types}\\
%\multicolumn{3}{c}{}\\
L ::= & & \textsc{{\tiny literal types:}}\\
&  \False \; | \; \True \; | \; {\it int} \; | \; {\it float} \; | \; {\it string} &\\
%\multicolumn{3}{c}{}\\
B ::= & & \textsc{{\tiny base types:}}\\
& \Boolean \, |\,  \Integer \, |\, \Number\, |\, \String &\\\\
%\multicolumn{3}{c}{}\\
%\multicolumn{3}{c}{}\\
V ::= & & \textsc{{\tiny value types:}}\\
& \;\; F \; | \; \Const \; F &\\
S ::= & & \textsc{second-level types:}\\
& \;\; P & \textit{tuple types}\\
& | \; S_{1} \sqcup S_{2} & \textit{unions of tuple types}\\
%\multicolumn{3}{c}{}\\
P ::= & & \textsc{{\tiny tuple types:}}\\
%& \;\; \Void & \textit{void type}\\
& | \; F{*} & \textit{variadic types}\\
& | \; F \times P & \textit{pair types}\\\\
\end{array}
$$
}
\caption{Type Language}
\label{fig:typelang}
\end{figure}

@@

In our example, variable $x$ would be bound to $\Integer \times \Integer \times \Nil{*}$ inside the {\bf then} branch, and bound
to $\Nil \times \String \times \Nil{*}$ inside the {\bf else}
branch. Now $q$ and $r$ project to $\Integer$ and $\Integer$
in the {\bf then} branch, and to $\Nil$ and $\String$ in the
{\bf else} branch.
